Argennon: A Scalable Smart Contract Platform

### 2022-03-28 tex/A.tex
Remove Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Re-add Argon chapter
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

\lstset{
    language=Java,
    morekeywords={contract, initialize, constructor, external, virtual, map, account, signature, overrides, is, string},
    basicstyle=\small\sffamily,
    columns=fullflexible,
%   keywordstyle=\bfseries,
    breaklines=true,
    showstringspaces=false,
    mathescape
}
%\binoppenalty=99999
%\relpenalty=99999
\newcommand{\note}[1] {
    \begin{tcolorbox}[colframe=white,colback=white]
        \emph{#1}
        %\emph{#1}\medskip
    \end{tcolorbox}
}

@@

    \chapter{The Argennon Virtual Machine}\label{ch:AVM}
    \input{sections/avm}

>>>>

%\binoppenalty=99999
%\relpenalty=99999
\newcommand{\note}[1] {
    \begin{tcolorbox}[colframe=white,colback=white]
        \emph{#1}
        %\emph{#1}\medskip
    \end{tcolorbox}
}

@@

    \chapter{The Argennon Smart Contract Execution Environment}\label{ch:AVM}
    \input{sections/avm}

### 2022-03-28 tex/sections/avm.tex
Save the draft of Ch1
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The Argennon\footnote{the classical pronunciation should be used:\textipa{/Ar"gen.non/}} Virtual Machine (AVM) is an
abstract computing machine for executing Argennon smart contracts. The
Argennon Virtual Machine knows nothing of the Argennon blockchain, transactions or any concept of gas usage, only of
Argennon identifier tries and the concept of execution sessions.

Execution sessions are separate sessions of executing smart contract's code by the Argennon Virtual Machine.
Every execution session starts with a single method invocation, but does not necessarily end with
that method invocation. During an execution session additional method invocations could be scheduled,
which will be executed after the initial method invocation completes.

An execution session may complete normally or abruptly. When an execution session completes abruptly, it will
not have any effects on the AVM's state.

\section{Data Types}\label{sec:data-types}

The Argennon Virtual Machine expects that all type checking is done prior to run time, typically by a compiler,
and does not have to be done by the Argennon Virtual Machine itself.

The Argennon Virtual Machine operates on two kinds of types: primitive types and identifier types. There are,
correspondingly, two kinds of values that can be stored in memory locations, passed as arguments, returned by
methods, and operated upon: primitive values and identifier values.

Values of primitive or identifier types need not be tagged or otherwise be inspectable to determine their types
at run time, or to be distinguished from values of other types. Instead, the instruction set of the Argennon
Virtual Machine distinguishes its operand types using instructions intended to operate on values of specific
types. For instance, \texttt{iadd64} assumes that its operands are two 64-bit signed integers or \texttt{delete}
assumes that its operand is an identifier type.

An identifier value in the Argennon Virtual Machine is a variable length array of bytes. Some instructions
that work with identifiers are able to determine the length of their identifier operands, while some
other instructions, for performance reasons, require the length to be specified.

The Argennon Virtual Machine does not have a fixed word size and uses \textbf{little-endian} byte order. Any
instruction of the AVM has its specific word size, and the addressable memory areas are byte addressable.

@@

In the Argennon Virtual Machine four distinct identifier types exist: \texttt{applicationID}, \texttt{accountID},
\texttt{methodID} and \texttt{chunkID}.

All these identifiers are \emph{prefix codes}, and hence can be represented by
\emph{prefix trees}\footnote{Also called tries.}. The Argennon has three primitive prefix trees:
\emph{applications, accounts} and \emph{local}. Any identifier in Argennon is a
prefix code built by one or more of these prefix trees:
\begin{itemize}
    \item \texttt{applicationID} is a prefix code built by \emph{applications} prefix tree.
    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An \texttt{accountID} can
    not be \texttt{0x0} or \texttt{0x1}.
    \item \texttt{methodID} is a composite prefix code built by concatenating an \texttt{applicationID} to
    a prefix code made by \emph{local} prefix tree: (\texttt{|} is the concatenating operator)
    \subitem \texttt{methodID = (applicationID|<local-prefix-code>)} .
    \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
    an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
    \subitem \texttt{chunkID = (applicationID|accountID)} or
    \subitem \texttt{chunkID = (applicationID|0x0|<local-prefix-code>)} or
    \subitem \texttt{chunkID = (applicationID|0x1|accountID|<local-prefix-code>)} .
\end{itemize}

All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
exmaple \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
\[
    (A^{(1)},A^{(2)},A^{(3)},\dots)\ ,
\]
where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)}<A^{(i+1)}\). A typical choice for \(\beta\)
could be \(2^8\).

@@

@@

\section{Architecture}\label{sec:arch}

\subsection{The \texttt{pc} Register}\label{subsec:the-pc-register}

The Argennon Virtual Machine always has a single thread of execution and exactly has one \texttt{pc} register.
When the Argennon Virtual Machine is executing a method, if that method is not native, the \texttt{pc} register
contains the address of the AVM instruction currently being executed. If the method currently being executed
is native, the value of the Argennon Virtual Machine's \texttt{pc} register is undefined.

\subsection{Call Stack Queue}\label{subsec:call-stack}

Every AVM execution session has a queue of call stacks. A call stack contains all the information that is needed for
restoring the AVM's state and continuing the execution after the method invocation completes.
This information is represented by a \texttt{CallInfo} struct:
\begin{verbatim}
CallInfo {
    methodID,
    pc,
    localFrame,
    operandStack
} .
\end{verbatim}

The \texttt{methodID} field is the unique identifier of a method's bytecode, which
is a composite prefix code built by concatenating an \texttt{applicationID} to
a local identifier.

Every method invocation has a corresponding \texttt{CallInfo} struct and when a method is invoked its
\texttt{CallInfo} struct is pushed onto the call stack. Hence, the top of the call stack always contains
the \texttt{CallInfo} struct of the current method. When a method invocation completes, its call
information is popped from the call stack.

An AVM execution session will continue as long as there is a non-empty call stack in the call stack queue. When the
execution of the current call stack finishes, the execution of the next call stack in the call stack queue starts.
It is possible that an AVM method invocation creates new call stacks. These newly created call stacks will be a part
of the current execution session and are added at the end of the call stack queue. A newly created call stack always
contains a single \texttt{CallInfo} struct. See Section~\ref{subsec:method-invocation} for more details.

\subsection{Run-Time Data Areas}\label{subsec:data-areas}

The Argennon Virtual Machine defines various run-time data areas that are used during an execution session,
Some of these data areas are persistent and are stored on the blockchain, Other data areas
are per execution session or per method invocation. These data areas are created when their context starts and
destroyed when their context ends.

The Argennon Virtual Machine has five run-time data areas:
\begin{itemize}
    \item Method Area
    \item Constant Area
    \item Local Frame
    \item Operand Stack
    \item Heap
\end{itemize}
All data areas except the operand stack, have their own address space. The operand stack is a
last-in-first-out (LIFO) stack and is not addressable. Every AVM instruction operates on its specific
data areas.

\subsubsection{Method Area}

The method area contains the byte-code of any method which can be called by a non-native method invocation
instruction. In the Argennon Virtual Machine every method has a unique identifier:
\texttt{methodID}, and the method area is a map from method identifiers to their byte codes.

Every smart contract is required to have a special \texttt{dispatcher} method. The \texttt{dispatcher} method is the
only method of a smart contract that can be called by other smart contracts and has a
predefined \texttt{methodID}: \texttt{(ApplicationID|0)}.

\note{The Argennon Virtual Machine does not have internal or external methods. All methods are considered internal
except the \texttt{dispatcher} method, which is the only external method of any smart contract.
This design enables the AVM to execute \emph{RESTful} smart contracts more efficiently.}

Instructions that modify the code area can only be run in privileged mode, which means they can only be executed
by the \emph{root} smart contract. As a result, the access of smart contracts to the code area is essentially
read-only.

\subsubsection{Constant Area}

Every smart contract has a single constant area which is stored in the AVM code area as a special type of
non-executable method. The constant area of a smart contract contains user
defined constants.

\note{Instructions that modify the constant area can only be run in privileged mode.}

\subsubsection{Local Frame}

A local frame is used to store methods parameters and local variables. A new frame is created each time a method
is invoked, and it is destroyed when its method invocation completes, whether the completion is normal or abrupt.

\subsubsection{Operand Stack}

Every time a local frame is created, a corresponding empty last-in-first-out (LIFO) stack is created too. The
Argennon Virtual Machine is a stack machine and its instructions take operands from the operand stack, operate on
them, and push the result back onto the operand stack. An operand stack is destroyed when its owner method
completes, whether that completion is normal or abrupt.

\subsubsection{Heap}

The heap of the Argennon Virtual Machine is a persistent memory area which stores \emph{memory chunks}. A Memory
chunk is a byte addressable piece of continuous memory which has a separate address space starting from 0. Each chunk
has a fixed size and different chunks need not be equally sized. Chunks are stored in the heap area and every chunk is
assigned a unique identifier: \texttt{chunkID}. As a result, the address of every memory location inside
the heap area can be considered as a pair: \texttt{(chunkID, offset)}.

A smart contract can read any chunk stored in the heap by having its \texttt{chunkID}. However, it can only modify
those chunks that was created by itself. In other words, every memory chunk in the AVM's heap has an owner. Only
the owner can modify a chunk while anyone can read that chunk. The owner of a chunk can be easily determined
by the \texttt{applicationID} part of the chunk identifier.

\note{The reason behind this type of access control design is the fact that smart contract code is usually
immutable. That means if a smart contract does not implement a getter mechanism for some parts of its internal
data, this functionality can never be added later, and despite the internal data is publicly available, there will
be no way for other smart contracts to use this data on-chain. This design eliminates the need for implementing
\emph{trivial} getters.}

The AVM Heap is able to save snapshots of its state and later restore them. This will enable the Argennon
Virtual Machine to have state reversion capability. The snapshot management of the heap area is done by
the following functions:
\begin{itemize}
    \item \texttt{Save()} saves a snapshot of the current state of the heap and pushes it onto the
    snapshot stack.
    \item \texttt{Restore()} pops the snapshot stored on the top of the snapshot stack and restores it. The
    current state of the heap will be lost.
    \item \texttt{Discard()} pops the snapshot stored on the top of the snapshot stack and discards it. The current
    state of the heap will not change.
\end{itemize}

\note{These functions are internal functions of the AVM. They are not instructions, and can not be called by smart
contracts.}

\section{Instruction Set Overview}\label{sec:instruction-set-overview}

An Argennon Virtual Machine instruction consists of a \textbf{one-byte} opcode specifying the operation to be
performed, followed by zero or more operands supplying arguments or data that are used by the operation.
The number and size of the operands are determined solely by the opcode.

\subsection{Method Invocation}\label{subsec:method-invocation}

The Argennon Virtual Machine has four types of method invocation instruction:
\begin{itemize}
    \item \texttt{invoke\_internal}: invokes a method without changing the context of method execution. In other
    words, the invoked method can only modify the state of the current smart contract.
    A smart contract can invoke any method by \texttt{invoke\_internal} instruction, even if that method is an
    internal method of another smart contract. Since the invoked method will always be executed in the context of
    the invoking smart contract, a smart contract will not be able to modify another smart contract's state by
    this instruction. This instruction facilitates code reuse and the usage of libraries.
    If the invoked method completes abruptly, the invoker method completes abruptly too.
    \item \texttt{invoke\_dispatcher}: invokes the \texttt{dispatcher} method of a smart contract and changes the
    context of method execution to that
    smart contract. If the invoked method completes abruptly, that will cause the invoker method to complete
    abruptly too.
    \item \texttt{i\_invoke\_dispatcher, i\_invoke\_internal}: independently invokes a method. These instructions
    are similar to their \texttt{invoke} counterparts, with the difference that when
    the invoked method completes abruptly, the invoker method can continue execution. When a method invoked by
    an \texttt{i\_invoke} instruction completes abruptly, all changes made
    by the invoked method to the AVM's state will be reverted and the instruction will act like
    a jump to the specified error handling code.
    \item \texttt{invoke\_native}: invokes a method that is not hosted by the Argennon Virtual Machine. By this
    instruction, high performance native methods of the hosting machine, such as GPU accelerated operations,
    could become available to AVM smart contracts. This instruction will not modify the AVM's state.
    \item \texttt{invoke\_later}: invokes the \texttt{dispatcher} method of another smart contract by creating a new
    call stack. The created call stack is added at the end of the call stack queue, and
    the context of method execution will change after the execution of that call stack starts. This instruction does
    not modify the \texttt{pc} register and the execution of the
    invoker will continue after this instruction. As a result, \texttt{invoke\_later} can not return
    any values to the caller and the return value of the called method (if any) will be discarded. If the invoked
    method completes abruptly, it will cause \textbf{all the call stack queue to
    complete abruptly.}
\end{itemize}

\note{The \texttt{invoke\_later} instruction can be used to call a smart contract while its reentracy lock is active.
Because the invocation uses a separate call stack, it will not cause the lock to throw an exception.}

Each time a method is invoked a new local frame
and operand stack is created. If the method was invoked by an \texttt{i\_invoke} instruction,
the \texttt{Save()} function of the AVM's heap is called before the execution of the method starts.

The Argennon Virtual Machine uses local frames to pass parameters on
method invocation. On method invocation, any parameters are passed in consecutive local variables stored in the
method's local frame starting from address 0. The invoker of a method writes the parameters in the local frame
of the invoked method using \texttt{arg} instructions.

\note{The AVM does not natively support polymorphism and virtual methods. A compiler could easily
generate appropriate code for implementing these features.}

\subsection{Exceptions}\label{subsec:exceptions}

An exception is thrown programmatically using the \texttt{athrow} instruction. Exceptions can also be thrown by
various Argennon Virtual Machine instructions if they detect an abnormal condition.

\note{Any AVM instruction may throw an exception. This happens when the execution cost of a transaction has exceeded
the declared maximum execution cost.}

\subsection{Method Invocation Completion}\label{subsec:method-invocation-completion}

A method invocation completes normally if that invocation does not cause an exception to be thrown, either
directly from the AVM or as a result of executing an explicit throw statement. If the invocation of the current
method completes normally and the invocation was not made by an \texttt{invoke\_later} instruction, then a value may be
returned to the invoking method. This occurs when the invoked method executes one of the return instructions, the
choice of which must be appropriate for the type of the value being returned (if any). Execution then continues
normally in the invoking method's local frame with the returned
value (if any) pushed onto the operand stack. If the method was invoked by an \texttt{invoke\_later} instruction,
the execution of the next call stack (if any) begins and the returned value
is discarded. When a method invocation that was made by
an \texttt{i\_invoke} instruction completes normally,
the \texttt{Heap.Discard()} function is called as a part of the return instruction.

A method invocation completes abruptly when an exception is thrown by one its instructions.
The \texttt{invoke\_internal} and \texttt{invoke\_dispatcher} instructions
always throw an exception when their invoked method completes abruptly. A
method invocation that completes abruptly never returns a value.

When a method completes normally, the call stack is used to restore the state of the invoker,
including its local frame and operand stack, with the \texttt{pc} register appropriately restored and incremented
to skip past the method invocation instruction. If the current call stack is empty the next call stack in the call
stack queue will be used for continuing the execution, and when there are no call stacks left in the queue the
current execution session ends.

When a method completes abruptly, the call stack is used to restore the \texttt{pc} register of the invoker. If the pc
register points to an \texttt{invoke} instruction, the current method completes abruptly as well. Otherwise, if
the \texttt{pc} register is pointing to an \texttt{i\_invoke} instruction, the \texttt{Heap.Restore()} function is
called to restore the state of the AVM's heap. Then, the
local frame and operand stack of the invoker is restored, and execution continues after branching to the error handling
code specified by the instruction's operand.

When an invocation which was made by an \texttt{invoke\_later} instruction completes abruptly, the
heap snapshot that was saved before the start of the
execution of the first call stack is restored. This means
that all changes made by the execution session to the AVM's heap will be completely reverted.

\subsection{Reentrancy Protection}\label{subsec:reentrancy}

@@

\subsection{Heap Allocation and Deallocation}\label{subsec:heap-allocation-instructions}

Heap allocation and deallocation in the Argennon Virtual Machine is done by the following instructions:

\begin{itemize}
    \item \texttt{alloc}: creates a new memory chunk of the specified size in the heap. The
    new chunk's identifier will be the concatenation of the \texttt{applicationID} of the creator of the chunk and
    the \texttt{id} operand of the instruction. The \texttt{id} operand must be the concatenation of
    an \texttt{accountID} and a prefix code
    generated by the \emph{local} prefix tree: \texttt{id = accountID|localID}. If the identifier is not valid or
    if it already exists in the heap, a catchable exception will be thrown.

    \item \texttt{delete}: deletes the chunk that its identifier is the concatenation of the current smart
    contract \texttt{applicationID} and the specified \texttt{id} operand. If the identifier is not valid or
    if the chunk was not found, a catchable exception will be thrown.
\end{itemize}

When a smart contract allocates a new memory chunk, the identifier of the new chunk is not generated by
the AVM. Instead, the smart contract can choose an identifier itself. This is a very important feature of
the AVM's heap, which allows smart contracts to use the AVM's heap as a dictionary (map) data structure.
Since the \texttt{chunkID} is a prefix code, any smart contract has its own identifier space, and a smart contract
can easily generate unique identifiers for its chunks.

The AVM requires the chosen identifier to be a prefix code compatible with
the \emph{accounts} prefix tree, or a code made by concatenating a code compatible with
the \emph{accounts} prefix tree to a code compatible with the \emph{local} prefix tree. This requirement
enables AVM to detect invalid identifiers, while smart contracts can easily associate data with account identifiers.
At the same time, A smart contract is able to create maps with costume keys by appending local identifiers
to the zero \texttt{accountID}. Zero \texttt{accountID} does not correspond to a valid account in the Argennon
blockchain.

@@

\section{The AVM Standard Library}\label{sec:the-avm-standard-library}

>>>>

The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
An Argennon application essentially is a http server whose state is kept in the Argennon blockchain and
its logic is described using an Argennon Standard Representation (ASR).

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations, one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other one is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. On the other hand, the low level language enables efficient
execution of applications that are not frequently used.

The state of an Argennon application is stored in byte addressable finite arrays of memory called
\emph{heap chunks}. An application can have several heap chunks with different sizes, and can remove or
resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
by its owner. In addition to heap chunk, every application also has an amount of non-persistent local memory for
storing temporary data.

The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
their http response and updating related heap
chunks. The AscEE executes requests sequentially \footnote{Actually requests are executed in parallel but by
performing data dependency analysis the result is guaranteed to be identical with sequential execution of
requests.} and each request has a separate \emph{execution session}. Execution sessions are separate sessions of
executing smart contract's code by the Argennon Smart Contract Execution Environment. The state of an execution
session will be
destroyed at the end of the session and only the state of heap chunks is preserved. If a session fails and does not
complete normally, it will not have any effect on the state of heap chunks.

During an execution session an application can make http requests to other applications. Those requests will not start
a new execution session and will be executed within the current session. In AscEE making a http request is similar to
a function invocation, because of that, we sometimes refer to it as an application
call.

The AscEE is designed based on \emph{optional
decoupling principle}. When an application makes a request to another application, optionally it can request to be
decoupled from the called application. That would mean the called application could not affect its caller's state
by reentrancy, or could not fail the session by using excessive resources or performing illegal operations.

@@

In Argennon a unique identifier is assigned to every application, heap chunk and account. Therefore, three distinct
identifier types exist: \texttt{appID}, \texttt{accountID}, and \texttt{chunkID}.
All these identifiers are \emph{prefix codes}, and hence can be represented by
\emph{prefix trees}\footnote{Also called tries.}.

Argennon has three primitive prefix trees:
\emph{applications, accounts} and \emph{local}.
All these trees are in base 256, with the maximum height
of 8. An Argennon identifier may be simple or compound. A simple identifier is generated using a single trie, while a
compound identifier is generated by concatenating prefix codes generated by two or more tries:

\begin{itemize}
    \item \texttt{appID} is a prefix code built by \emph{applications} prefix tree. An application ID cannot
    be \texttt{0x0}.

    \item \texttt{accountID} is a prefix code built by \emph{accounts} prefix tree. An account ID cannot
    be \texttt{0x0} or \texttt{0x1}.

    \item \texttt{chunkID} is a composite prefix code built by concatenating an \texttt{applicationID} to
    an \texttt{accountID} to a prefix code made by \emph{local} prefix tree:
    \subitem \texttt{chunkID = (applicationID|accountID|<local-prefix-code>)} .
\end{itemize}

All Argennon prefix trees have an equal branching factor \(\beta\). Therefore, we can represent an Argennon
prefix tree as a sequence of fractional numbers\footnote{It's possible to have \(a_i=0\). For
exmaple \(A^{(4)}=(0.2000)_{10}\) is correct.} in base \(\beta\):
\[
    (A^{(1)},A^{(2)},A^{(3)},\dots)\ ,
\]
where \(A^{(i)}=(0.a_{1}a_{2}\dots a_{i})_\beta\), and we have \(A^{(i)} \leq A^{(i+1)}\). A typical choice for
\(\beta\) could be \(2^8\).

@@

\section{Heap Chunks}\label{heap}

The persistent data of an Argennon application is stored in heap chunks. An application may have several heap chunks
with different sizes, and can remove or
resize its chunks or allocate new chunks. Every chunk belongs to exactly one application. Only the owner application can
modify a chunk but there is no restrictions for reading a chunk.

When an application allocates a new heap chunk, the identifier of the new chunk is not generated by
the AscEE. Instead, the application can choose an identifier itself \footnote{provided it has a correct format}. This
is an important feature of the AscEE's heap, which allows applications to use the AscEE's heap as a dictionary (map)
data structure.
Since the \texttt{chunkID} is a prefix code, any application has its own identifier space, and an application
can easily generate unique identifiers for its chunks.

A heap chunk can be considered as a continuous
array of bytes. Every chunk has a size: \texttt{chunkSize} and a size upper bound: \texttt{sizeUpperBound}. The value of
\texttt{chunkSize} can be determined uniquely at
the start of
every execution session, and it may be updated during the session like a normal memory location. On the other hand,
the value of \texttt{sizeUpperBound} is constant for every block of the blockchain and is proposed by the block
proposer.

The address space of a chunk starts from zero and only offsets lower than \texttt{sizeUpperBound} are valid. Trying to
access any offset higher than this value will result in a revert for the application.

\subsection{Access Blocks}\label{subsec:access-blocks}

Memory locations inside a chunk can only be accessed through access blocks. An access block is defined on a chunk
and has an offset and a size and determines accessible memory locations inside a chunk. Multiple access blocks can
be defined on a chunk, but they must be non overlapping. Access blocks are byte addressable and can have different
access types:

\begin{itemize}
    \item \texttt{read\_only}: only allows read and check operations.
    \item \texttt{writable} memory locations inside this access block can not be modified.
    \item \texttt{check\_only}: only allows check operations. These operations query the persistence
    status of memory locations.
    \item \texttt{additive}: only allows addition operations without overflow checking. Note that the content of these
    access blocks cannot be read.
\end{itemize}

Chunks are intended for simplifying proof checking of the data stored in the Argennon cloud and access blocks are
required for better parallelization of request execution. An application should put the data it predicts is needed for
validating a block in the same chunk and put the data it predicts will be needed in a single execution session in
the same access block.

\subsection{Chunk Resizing}\label{subsec:ch-resize}

The value of \texttt{chunkSize} at the end of the execution session will determine if a memory location at an
offset is persistent or not: Offsets lower than the chunk size are persistent, and higher offsets are not.
Non-persistent locations will be re-initialized with zero, at the start of every execution session.

The value of chunkSize can be modified during an execution session. However, the new size values can only be
increasing or decreasing. More precisely, if a request declares that it wants to expand (shrink) a chunk, it can only
increase (decrease) the value of chunkSize and any specified value for chunkSize during the execution session, needs
to be greater (smaller) than the previous value of the chunk size. Any request that wants to expand (shrink) a chunk
needs to specify a max size (min size). The value of chunkSize can not be set higher (lower) than this value.

Usually an application should not have any assumption about the content of memory locations that are outside the chunk.
While these locations are zero initialized at the start of every execution session, it should be noted that multiple
invocations of an application may occur in a single execution session, and if one of them modifies a location outside
the chunk, the changes can be seen by next invocations.

There is no way for an application to query the chunk capacity. As a result, in the view of an application, accessing
offsets higher than chunkSize results in undefined behaviour, while the behaviour is well-defined for validators.
This enables validators to determine the validity of an offset at the start of the block validation in a parallelized
preprocessing phase without executing requests.

While an application can use \texttt{chunkSize} to determine if an offset is persistent or not, that is not
considered a good practice. Reading \texttt{chunkSize} decreases transaction parallelization, and should be avoided.
Instead, applications should use a built-in AscEE's function for checking the persistence status of memory addresses.

An application can load a chunk with a valid prefix identifier even if that chunk does not exist. For a non-existent
chunk the value of \texttt{chunkSize} is always zero.

\section{Resource Management}\label{sec:res-man}

Completing an execution session needs computational resources. The amount of resources used by an execution session
needs to be monitored, otherwise a spammer would be able to easily exhaust the resources of the execution
environment. Resource usage can be measured per session or per application call.

The AscEE has two type of execution sessions: optimistic and monitored. Resource usage of an optimistic session
is always measured per session and default pre-defined resource caps are used. An optimistic session must always
complete normally. On the other hand in a monitored session some resources are measured per application and some
caps are determined dynamically by the request. A monitored session is allowed to either complete normally or fail.

Obviously resource monitoring is easier for optimistic sessions, and they can be executed more efficiently. However, if
an optimistic session violates its resource caps, determining the point of failure requires precise measurement of
used resources. Note that for implementing optional decoupling principle we need to determine the exact application
which fails in the call chain. For example assume that an optimistic session containing an application call violates a
1ms execution time cap. If the call happens exactly after 1ms a small fluctuation in measuring the execution time can
change the point of failure between the called and the main application. On the other hand this is not an issue for
monitored session because resource measurements and caps are defined per application.

The following resources are defined and monitored for AscEE execution sessions:

execution time:
The amount of cpu time that is required for executing a session or an application. The execution time of a session
is measured in \emph{AscEE clocks}. One AscEE clock is defined as 1/1000 of the amount
of cpu time needed for executing a predefined standard application which is used for benchmarking a host's performance.

local memory: any memory usage of an application that is not part of the heap chunks and is not part of another resource
will be considered as local memory usage. This resource is measured in bytes.

heap access list: every session can only access heap locations that are declared in its access list. In addition,
resizing
heap chunks can only be done in the range of the pre-declared lower bound and upper bound. those predeclared
locations of heap chunks that are

applications list: a session can only make requests to applications that are declared in its application list.

call depth: the number of nested application calls.

number of deferred calls
number of session locks
number of virtual signatures

@@

\section{Reentrancy Protection}\label{sec:reentrancy}

@@

\section{Deferred Calls}\label{sec:deferred-calls}

@@

\section{The Argennon Standard Library}\label{sec:asl}
